Höglinger et al. 2017

### 2025-02-20 IndividualStudies/DH_2017.qmd
Spent all day learning how to make my heatmap look less bad

## intitialize libraries
library(dplyr)
library(tidyr)
library(plotly)
library(ggplot2)
library(grid)
library(readr)
library(rmarkdown)
library(emojifont)
library(gt)
library(crosstalk)
library(DT)
library(stringr)
library(pheatmap)
library(here)

@@

DH_2017_data <- read.csv("DataTables/DH_PNAS_2017_download.csv") |>
  mutate(gene_name = str_extract(Description, "GN=([^\\s]+)") |> str_remove("GN=")) |>
  mutate(Species = str_extract(Description, "OS=([^=]+?)\\sGN=") |> str_remove("OS=") |> str_remove("GN=")) |>
  filter(Species == "Homo sapiens ") |>
  mutate(PSM_Sph = (PSM_Sph1 + PSM_Sph2) / 2,
         PSM_FA = (PSM_FA1 + PSM_FA2) / 2,
         PSM_DAG = (PSM_DAG1 + PSM_DAG2) /2) |>
  select(gene_name, PSM_Sph, PSM_FA, PSM_DAG) |>

  glimpse()

```

@@

#### Heat map

```{r, echo=FALSE, warning=FALSE, fig.height=15}

DH_2017_data_long <- read.csv("DataTables/DH_PNAS_2017_long.csv") |>
  mutate(PSM = value, sample = name) |>
  group_by(gene_name, sample) |>
  summarise(PSM = mean(PSM, na.rm = TRUE), .groups = "drop") |>
  distinct()

DH_2017_data_long$sample <- factor(pull(DH_2017_data_long, sample), ordered = TRUE, levels = c("Sph", "FA", "DAG"))

heatmap <- DH_2017_data_long |>
  ggplot(aes(y = gene_name, x = sample)) +
  theme(panel.border = element_rect(color = NA, fill = "grey25")) +
  geom_tile(aes(fill = PSM, text = paste0("Gene name: ", gene_name, '\n', "Lipid Probe:", sample, "\n", "PSM: ", PSM))) +
  scale_fill_gradientn(colors = c("grey95", "purple", "orange"),
  values = scales::rescale(c(0, 10, 25))) +
  ylab("Gene") +
  xlab("Lipid Probe") +
  theme_minimal() +
  theme(text = element_text(family = "serif")) +
  theme(legend.title = element_text(size = 12, face = "bold")) +
  theme(axis.title = element_text(size = 12, face = "bold")) +
  theme(legend.text = element_text(size = 12))

m <- list(l=50, r = 50, b= 100, pad = 4)

ggplotly(heatmap,
        tooltip = c("text", dynamicTicks = TRUE),
        hovermode = "closest",
        autosize = T,
        margin = m) |>
  layout(showlegend = FALSE)

```

@@

>>>>

## intitialize libraries
library(dplyr)
library(tidyr)
library(plotly)
library(ggplot2)
library(grid)
library(readr)
library(rmarkdown)
library(emojifont)
library(gt)
library(crosstalk)
library(DT)
library(stringr)
library(pheatmap)
library(here)
library(heatmaply)
library(ggvenn)
library(scales)

@@

DH_2017_data <- read.csv(paste0(here(),"/IndividualStudies/DataTables/DH_PNAS_2017_download.csv")) |>
  mutate(gene_name = str_extract(Description, "GN=([^\\s]+)") |> str_remove("GN=")) |>
  mutate(Species = str_extract(Description, "OS=([^=]+?)\\sGN=") |> str_remove("OS=") |> str_remove("GN=")) |>
  filter(Species == "Homo sapiens ") |>
  mutate(PSM_Sph = (PSM_Sph1 + PSM_Sph2) / 2,
         PSM_FA = (PSM_FA1 + PSM_FA2) / 2,
         PSM_DAG = (PSM_DAG1 + PSM_DAG2) /2) |>
  select(gene_name, PSM_Sph, PSM_FA, PSM_DAG) |>
  arrange(gene_name) |>
  group_by(gene_name) |>
  summarise(PSM_DAG = mean(PSM_DAG),
            PSM_FA = mean(PSM_FA),
            PSM_Sph = mean(PSM_Sph))
```

@@

#### Heat map
:::{.border}
```{r, echo=FALSE, fig.height=10}

DH_2017_data_long <- DH_2017_data |>
  pivot_longer(cols = starts_with("PSM_"),
              names_to = "variable",
              names_prefix = "PSM_",
              values_to = "value") |>
  mutate(name = gene_name) |>
  select(name, variable, value)
  # mutate(OriginalValues = value) |>
  # mutate(value = case_when(value < 20 ~ value,
  #                           value >= 20 ~ 20))

DH_2017_data_long$variable <- factor(pull(DH_2017_data_long, variable), ordered = TRUE, levels = c("Sph", "FA", "DAG"))

# Create the heatmap
heatmaply(long_data = DH_2017_data_long,
          xlab = "Probe",
          ylab = "Gene name",
          label_names = c("Gene name", "Probe", "PSM count"),
          scale_fill_gradient_fun = ggplot2::scale_fill_gradientn(colours = c("grey95", "purple", "orange", "orange", "orange", "orange", "orange", "red", "red","red"), breaks = c(0, 5, 20, 200)),
          showticklabels = F)

```
:::

<hr>

@@

:::{.border}
```{r, echo=FALSE, warning=FALSE, height = 5}

# Ensure the data is in the correct format
DH_2017_data <- as.data.frame(DH_2017_data)
rownames(DH_2017_data) <- DH_2017_data[,1]
DH_2017_data[,1] <- NULL

Sph_proteins <- DH_2017_data |>
  filter(PSM_Sph > 0) |>
  select(PSM_Sph)

FA_proteins <- DH_2017_data |>
  filter(PSM_FA > 0) |>
  select(PSM_FA)

DAG_proteins <- DH_2017_data |>
  filter(PSM_DAG > 0) |>
  select(PSM_DAG)

combined_venn_proteins <- list("Sph proteins" = rownames(Sph_proteins), "FA proteins" = rownames(FA_proteins), "DAG proteins" = rownames(DAG_proteins))

ggvenn(combined_venn_proteins)

```
:::
